### 2021-03-24 paper/00_main.tex
Added paper template.

% \documentclass[review]{cvpr}
\documentclass[final]{cvpr}

\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[pagebackref=true,breaklinks=true,colorlinks,bookmarks=false]{hyperref}

\def\cvprPaperID{****} % *** Enter the CVPR Paper ID here
\def\confYear{CVPR 2021}
% \setcounter{page}{4321} % For final version only

\begin{document}

%%%%%%%%% TITLE
\title{AI604 term project paper template}

\author{Except for the proposal, do not provide student information (e.g., name and student ID)
}

\maketitle

\begin{abstract}
Abstract is optional. This will not affect the final score.
\end{abstract}

\section{Guideline: Proposal}

Feel free to describe your idea. We expect that the proposal includes \textit{the definition of the task}, \textit{dataset}, and \textit{tentative plan for the project}.

Please note that you can change the project topic after the proposal. However, we strongly discourage you from changing the topic after progress 1.

\section{Guideline: Progress 1 / 2}

After the proposal, we will check the progress of your project two times. You should submit the progress reports, and the reports should contain the contents of the proposal, \textit{related work} (and \textit{previous work} if exist), \textit{method}, and \textit{experiment}. It is okay to omit the result.

Other students will review your progress reports, and it is in a fully \textit{double-blind} manner. Each team will receive three or four reviews, and the review scores will affect your final score. For fair evaluation, students who have the same advisor do not review each others' work, and also \textbf{you will judge the reviews' review!} In the case of Progress 1, every team should submit \textit{author response}. The length of the response is up to one page. We provide the template for the author response; please check \texttt{01\_rebuttal.tex}.

\section{Guideline: Final}

The final report should contain the contents of the proposal and the progress reports, \textit{result}, and \textit{etc}. The length of the paper is up to nine pages except for \textit{reference}. Reference is optional, but we strongly recommend you to add the reference.

There will be a project presentation session at the end of the semester. Further instructions will be announced later.

\section{Etc}

If you have a question, send us an e-mail or a message to the Kakao open chat room.

{\small
\bibliographystyle{ieee_fullname}
\bibliography{reference}
}

\end{document}