### 2021-04-25 paper/progress-1/01_rebuttal.tex
Added progress report 1.

\documentclass[rebuttal]{cvpr}

\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[pagebackref=true,breaklinks=true,colorlinks,bookmarks=false]{hyperref}

%%%%%%%%% PAPER ID  - PLEASE UPDATE
\def\cvprPaperID{0000} % Do not enter team number here
%\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

\def\cvprPaperID{0000} % Do not enter team number here
\def\confYear{AI604 2021 spring}

\begin{document}

%%%%%%%%% TITLE - PLEASE UPDATE
\title{AI604 term project rebuttal / author response}  % **** Enter the paper title here

\maketitle
\thispagestyle{empty}

%%%%%%%%% BODY TEXT - ENTER YOUR RESPONSE BELOW
\section{Introduction}

After receiving paper reviews, authors should submit a rebuttal to address the reviewers' comments, which will be limited to a {\bf one page} PDF file.

Note that the author rebuttal is not visible to the reviewers. In other words, only TAs and the professor will read the rebuttal.

\subsection{Response length}

Author responses must be no longer than 1 page in length except for \textbf{References} and \textbf{Additional question}.

\subsection{Note}

Feel free to write the author response, and we recommend you not to spend your time so much on this. We aim to provide you a chance for improve your work by answering to the reviews.

%-------------------------------------------------------------------------
%\newpage

\section{Additional question}
Each team should answer the below questions for evaluating their reviewers.

\subsection{Question 1}
\begin{itemize}
    % 리뷰어 1 2 3 은 우리의 프로젝트 내용을 이해했다.
    \item The reviewer 1 understood our main idea and contributions. (Yes / Neutral / No)
    \item The reviewer 2 understood our main idea and contributions. (Yes / Neutral / No)
    \item ...
\end{itemize}

\subsection{Question 2}
\begin{itemize}
    % 리뷰어 1 2 3 은 우리에게 건설적인 피드백을 제공했다.
    \item The reviewer 1 provided constructive feedback or advice. (Yes / Neutral / No)
    \item The reviewer 2 provided constructive feedback or advice. (Yes / Neutral / No)
    \item ...
\end{itemize}

\subsection{Question 3}
\begin{itemize}
    % 기타 하고 싶은 말
    \item Other comments for TAs and the professor.
\end{itemize}

{\small
\bibliographystyle{ieee_fullname}
\bibliography{reference}
}

\end{document}